Boosting RAG Efficiency with dyRAG: Dynamic K-Selection for Optimal Retrieval

### 2025-07-18 Paper/JAIR_Example_Template/JAIR_RAG_Paper.tex
All changes
Last submitted paper is in folder submission3

	%Version 3 October 2023
% See section 11 of the User Manual for version history
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing

%%=======================================================%%
%% to print line numbers in the margin use lineno option %%
%%=======================================================%%

%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%======================================================%%
%% to compile with pdflatex/xelatex use pdflatex option %%
%%======================================================%%

%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%Note: the following reference styles support Namedate and Numbered referencing. By default the style follows the most common style. To switch between the options you can add or remove “Numbered” in the optional parenthesis. 
%%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst%

%%\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals
%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
\documentclass[sn-mathphys-num]{sn-jnl}% Math and Physical Sciences Numbered Reference Style 
%%\documentclass[sn-mathphys-ay]{sn-jnl}% Math and Physical Sciences Author Year Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style 
%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style

%%%% Standard Packages
%%<additional latex packages if required can be included here>

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
\usepackage{subcaption}

%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\begin{document}

	\title[Article Title]{dyRAG: Dynamic Retrieval Augmentation for Adaptive RAG Performance}

	%%=============================================================%%
	%% GivenName	-> \fnm{Joergen W.}
	%% Particle	-> \spfx{van der} -> surname prefix
	%% FamilyName	-> \sur{Ploeg}
	%% Suffix	-> \sfx{IV}
	%% \author*[1,2]{\fnm{Joergen W.} \spfx{van der} \sur{Ploeg} 
	%%  \sfx{IV}}\email{iauthor@gmail.com}
	%%=============================================================%%

	\author*[1]{\fnm{Zoubida Asmaa} \sur{Boudjenane }}\email{asmaa.boudj04@gmail.com}

	\author[2]{\fnm{Mohammed} \sur{Salem}}\email{salem@univ-mascara.dz}
	\equalcont{These authors contributed equally to this work.}

	%\author[1,2]{\fnm{Third} \sur{Author}}\email{iiiauthor@gmail.com}
	%\equalcont{These authors contributed equally to this work.}

	\affil*[1]{\orgdiv{Computer Science Department}, \orgname{University of Mascara}, \orgaddress{\street{Street}, \city{Mascara}, \postcode{29000}, \state{}, \country{Algeria}}}

	\affil*[2]{\orgdiv{LISYS Lab}, \orgname{University of Mascara}, \orgaddress{\street{Street}, \city{City}, \postcode{100190}, \state{State}, \country{Algeria}}}

	%\affil[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{610101}, \state{State}, \country{Country}}}

	%%==================================%%
	%% Sample for unstructured abstract %%
	%%==================================%%

	\abstract{Retrieval-Augmented Generation (RAG) enhances language model responses by incorporating external knowledge. However, its effectiveness heavily depends on the quality of the retrieved documents. Using a fixed number of retrieved documents K may fail  to adapt to varying query complexity, often leading to suboptimal retrieval either including irrelevant documents or missing crucial ones. To address this issue, we propose dyRAG, a hybrid method that dynamically adjusts K based on query characteristics while maintaining computational efficiency. Our method improves retrieval performance by maximizing relevant information and minimizing noise. Experimental results show that it outperforms fixed-K retrieval, offering a more effective solution for optimizing RAG systems. Compared to traditional methods, it dynamically adjusts based on the query and dataset characteristics, offering greater adaptability across various contexts.}

	%%\pacs[JEL Classification]{D8, H51}

	%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

	\maketitle

	\section{Introduction}\label{sec1}

	Large Language Models (LLMs) have emerged as state-of-the-art artificial intelligence systems that excel at understanding and generating human-like text\cite{Radford2019}, Built on deep learning architectures, particularly transformers \cite{vaswani2017attention},these models showcase impressive adaptability across diverse tasks, such as machine translation\cite{Gu2018SearchEngine}, text generation\cite{liang2024controllabletextgenerationlarge} and question answering\cite{Zhang2023}. Notable examples of LLMs include OpenAI GPT series\cite{brown2020language}, Google BERT\cite{devlin2018bert} and T5 \cite{raffel2020exploring}. Despite their impressive capabilities, LLMs suffer from hallucination\cite{Radford2019}, often generating confident but incorrect information. This is a critical challenge for knowledge-intensive tasks requiring high accuracy\cite{Radford2019}. Additionally, their static nature being trained on fixed datasets\cite{Radford2019} limits their ability to update knowledge dynamically, making them prone to outdated or inconsistent responses, especially for time-sensitive information.\cite{Mousavi2025}.

	Researchers have explored various approaches to address this limitation, including transfer learning\cite{zhuang2020comprehensive}, fine-tuning \cite{Howard2018} which adapt pre-trained models to specific tasks. Despite their strengths, they are computationally expensive and impractical for real-time knowledge updates. A promising alternative is Retrieval-Augmented Generation (RAG) \cite{lewis2020retrieval}, which integrates real-time retrieval from external knowledge sources with generative models, enables LLMs to access up-to-date information without retraining,(RAG) faces several limitations. For instance, The performance of RAG heavily depends on the quality of the retriever \cite{karpukhin2020dpr}, as irrelevant or noisy documents can degrade the generator output. Additionally, a critical challenge is determining the optimal number of retrieved passages K to feed into the generator\cite{lewis2020retrieval}. Often resulting in either insufficient or excessive information being retrieved, reduce coherence and increase computational overhead. This trade-off makes K a key factor in balancing performance and efficiency.

@@

	\section{Overview of Retrieval-Augmented Generation }\label{sec3}
	Retrieval-Augmented Generation (RAG) is a method designed to enhance the performance of large language models (LLMs) by integrating external, reliable knowledge sources into their response generation process\cite{awsRAG}. RAG combines search capabilities with LLM prompting. The process involves feeding both the userss query and the relevant information retrieved by a search algorithm into the LLMs prompt\cite{ilin2023advancedrag}, enabling the model to generate answers based on the provided context.
	gupta20241
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.9\textwidth]{rag.JPG}
		\caption{RAG system component\cite{gupta20241}. }\label{rag.PNG}
	\end{figure}
	As illustrated in figure \ref{rag.PNG}, the RAG system is built around two main components: the retriever and the generator. The retriever's role is to find relevant information from a pre-built data store, while the generator utilizes this retrieved data to produce coherent and contextually appropriate content. The overall RAG process operates as follows :
	\subsection{Retrieval in RAG Systems}\label{subsec2}

@@

	The query \( q \) represents the user's state at a specific time step, derived from their interaction history. In the MoL (Mixture of Logits) framework,\( q \)is transformed into \( Pq \)
	embeddings through a multi-layer perceptron (MLP).
	\subsection{Impact of Hyperparameters}
	For fair comparison, we maintained consistent architectural choices and training conditions across all experiments, We conducted an extensive hyperparameter analysis comparing both approaches (Hybrid+SAS and MoL+SAS) across different architectural configurations. All experiments were Implemented in TensorFlow and trained on a Google Colab environment with a T4 GPU. We use the Adam optimizer with a learning rate of 0.001. For the hybrid algorithm, we initialize the threshold (Tinit) to 0.3 and adaptively adjust it during training. We discuss detailed hyperparameter settings in table \ref{tab:100k_movies}, table \ref{tab:1m_movies}

@@

	\begin{table}[ht]
		\centering
		\large
		\renewcommand{\arraystretch}{1.7} % Increase row spacing
		\resizebox{\textwidth}{!}{  % Scale table to fit width
			\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
				\hline
				\textbf{Model} & \textbf{\shortstack{Max \\ Sequence Length}} & 
				\textbf{\shortstack{Embedding \\ Dimension}} & 
				\textbf{\shortstack{Number \\ of Heads}} & 
				\textbf{\shortstack{Feedforward \\ Dimension}} & 
				\textbf{\shortstack{Batch \\ Size}} & 
				\textbf{Epochs} & 
				\textbf{\shortstack{Val \\ Loss}} & 
				\textbf{\shortstack{Val \\ Accuracy}} \\
				\hline
				Hybrid+SAS & 50  & 128  & 2 & 128  & 128 & 12 & 5.3036 & 0.1584  \\ \hline
				MoL+SAS    & 50  & 128  & 2 & 128  & 128 & 12 & 5.3152 & 0.1582  \\ \hline
				Hybrid+SAS & 128 & 256  & 4 & 256  & 128 & 10 & 3.6364 & 0.4301  \\ \hline
				MoL+SAS    & 128 & 256  & 4 & 256  & 128 & 10 & 3.6374 & 0.4299 \\\hline
				Hybrid+SAS & 512 & 512  & 4 & 512  & 128 & 10 & 1.2808 & {\bf 0.8120} \\ \hline
				MoL+SAS    & 512 & 512  & 4 & 512  & 128 & 10 & 1.3050 & 0.8016 \\\hline
			\end{tabular}
		}
		\caption{Results on 100kMovies dataset}
		\label{tab:100k_movies}
	\end{table}

	\begin{table}[ht]
		\centering
		\large
		\renewcommand{\arraystretch}{1.7} % Increase row spacing
		\resizebox{\textwidth}{!}{  % Scale table to fit width
			\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
				\hline
				\textbf{Model} & \textbf{\shortstack{Max \\ Sequence Length}} & 
				\textbf{\shortstack{Embedding \\ Dimension}} & 
				\textbf{\shortstack{Number \\ of Heads}} & 
				\textbf{\shortstack{Feedforward \\ Dimension}} & 
				\textbf{\shortstack{Batch \\ Size}} & 
				\textbf{Epochs} & 
				\textbf{\shortstack{Val \\ Loss}} & 
				\textbf{\shortstack{Val \\ Accuracy}} \\
				\hline
				Hybrid+SAS & 50  & 128  & 2 & 128  & 128 & 10 & 4.5721 & 0.1530  \\ \hline
				MoL+SAS    & 50  & 128  & 2 & 128  & 128 & 10 & 4.5733 & 0.1526  \\ \hline
				Hybrid+SAS & 128 & 256  & 4 & 256  & 128 & 10 & 3.4152 & 0.3680  \\ \hline
				MoL+SAS    & 128 & 256  & 4 & 256  & 128 & 10 & 3.4671 & 0.3627  \\ \hline
				Hybrid+SAS & 512 & 512  & 4 & 512  & 128 & 10 & 1.0275 & {\bf 0.8350} \\ \hline
				MoL+SAS    & 512 & 512  & 4 & 512  & 128 & 10 & 1.0350 & 0.8276  \\ \hline

			\end{tabular}
		}
		\caption{Results on 1M Movies dataset}
		\label{tab:1m_movies}
	\end{table}
	\newpage
	Both approaches demonstrate significant performance improvements as model capacity increases, with larger configurations consistently delivering better results. For instance, when increasing the model's capacity such as expanding the (Max Sequence Length: 512, Embedding Dimension: 512, Number of Heads: 4, Feed-Forward Dimension: 512) the hybrid algorithm combined with SASRec (Hybrid+SAS) achieves notable gains, particularly in the most resource-intensive setup. On the ML-100K dataset, Hybrid+SAS reaches a score of\textbf{0.8120} compared to the baseline's \textbf{0.8016}, while on ML-1M, it achieves \textbf{0.8350} versus \textbf{0.8276} The ML-1M dataset generally benefits more from increased capacity, with the performance gap between datasets narrowing as the model scales. While smaller configurations provide a balance of efficiency and performance, the largest configuration, despite its higher computational demands, yields the best results, making it suitable for scenarios where resources are not a constraint. Both approaches show similar benefits from scaling, but Hybrid+SAS maintains a consistent edge in performance.

@@

	\newpage

	%%===================================================%%
	%% For presentation purpose, we have included        %%
	%% \bigskip command. Please ignore this.             %%
	%%===================================================%%

	%%===========================================================================================%%
	%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
	%% system, please include the references within the manuscript file itself. You may do this  %%
	%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
	%% file, and delete the associated \verb+\bibliography+ commands.                            %%
	%%===========================================================================================%%

	\bibliography{sn-bibliography}% common bib file
	%% if required, the content of .bbl file can be included here once bbl is generated
	%%\input sn-article.bbl

\end{document}

>>>>

%% This is a JAIR Example File Compiled by Nicholas Mattei (nsmattei@tulane.edu) 
%% and Odd Erik Gundersen (odderik@ntnu.no)
%% and Mykel Kochenderfer (mykel@stanford.edu)
%% March 29, 2025
%%
%% This file is based off the ACM Latex Template https://www.acm.org/publications/proceedings-template
%% Revision 2.12 (12/28/2024)
%% 
%% Please see https://www.jair.org/index.php/jair for more information and submission instructions.
%%

%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{jair}.
%%
\documentclass[review]{jair}

\setcopyright{cc}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{10.1613/jair.1.xxxxx}

%%
\JAIRAE{Insert JAIR AE Name}
\JAIRTrack{Insert JAIR Track Name Here}
\acmVolume{4}
\acmArticle{111}
\acmMonth{8}
\acmYear{2025}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\usepackage{graphicx}%
\usepackage{multirow}%
%\usepackage{amsfonts, amsmath,amssymb}%
%%\usepackage{amsthm}%
%%\usepackage{mathrsfs}%
%\usepackage[title]{appendix}%
%\usepackage{xcolor}%
%\usepackage{textcomp}%
%\usepackage{manyfoot}%
%\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
\usepackage{subcaption}
\usepackage{booktabs} % For professional-quality rules
\usepackage{makecell} % For multiline headers
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
%\title{JAIR Example Template}
\title[dyRAG.Dynamic K-Selection]{Boosting RAG Efficiency with dyRAG: Dynamic K-Selection for Optimal Retrieval}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research and/or corresponding author.
\author{Zoubida Asmaa Boudjenane}

\email{asmaa.boudj04@gmail.com}
\orcid{0000-0002-2037-3694}
% \orcid{https://orcid.org/0000-0002-2037-3694} % This works too if you want to see the full URL
\affiliation{%
  \institution{Computer Science Department, University of Mascara}
  \city{Mascara}
  \state{Mascara}
  \country{Algeria}
}

\author{Mohammed Salem}
\orcid{0000-0002-2037-3694}
\authornote{Corresponding Author.}
\email{salem@univ-mascara.dz}
\affiliation{%
  \institution{LISYS Laboratory, University of Mascara}
  \city{Mascara}
  \state{Mascara}
  \country{Algeria}}

%\author{Holger H. Hoos}
%\orcid{0000-0003-0629-0099}
%\email{hoos@cs.ubc.ca}
%\affiliation{%
%  \institution{University of British Columbia}
%  \city{Vancouver}
%  \state{British Columbia}
%  \country{Canada}
%}

%\author{Kevin Leyton-Brown}
%\orcid{0000-0002-7644-5327}
%\email{kevinlb@cs.ubc.ca}
%\affiliation{%
%  \institution{University of British Columbia}
%  \city{Vancouver}
%  \state{British Columbia}
%  \country{Canada}
%}

%\author{Ben Trovato}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Boudjenane and Salem.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
{\bf Background:} 
    Retrieval-Augmented Generation (RAG) enhances language model responses by incorporating external knowledge. However, its effectiveness heavily depends on the quality of the retrieved documents. Using a fixed number of retrieved documents (K) may fail to adapt to varying query complexity, often leading to suboptimal retrieval—either including irrelevant documents or missing crucial ones.

    {\bf Objectives:}
   To address this issue, we propose dyRAG, a hybrid method that dynamically adjusts K based on query characteristics while maintaining computational efficiency. Our goal is to improve retrieval performance by maximizing relevant information and minimizing noise.

    {\bf Methods:}
   We introduce dyRAG, which dynamically adjusts the number of retrieved documents (K) in response to query complexity and dataset characteristics. The method balances precision and computational efficiency to optimize RAG systems.

    {\bf Results:}
    Experimental results show that dyRAG outperforms fixed-K retrieval, offering a more effective solution for optimizing RAG systems. It demonstrates greater adaptability across various contexts compared to traditional methods.

    {\bf Conclusions:}
    dyRAG provides a dynamic and efficient approach to document retrieval in RAG systems, improving performance by tailoring retrieval to query-specific needs. This advancement enhances the flexibility and effectiveness of RAG in diverse applications.
\end{abstract}

%\begin{abstract}
%      A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

%% JAIR Note: 
%% Do not include ACM CCS Concepts or Keywords

%% To be updated by authors.
\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
Large Language Models (LLMs) have emerged as state-of-the-art artificial intelligence systems that excel at understanding and generating human-like text\cite{Radford2019}, Built on deep learning architectures, particularly transformers \cite{vaswani2017attention},these models showcase impressive adaptability across diverse tasks, such as machine translation\cite{Gu2018SearchEngine}, text generation\cite{liang2024controllabletextgenerationlarge} and question answering\cite{Zhang2023}. Notable examples of LLMs include OpenAI GPT series\cite{brown2020language}, Google BERT\cite{devlin2018bert} and T5 \cite{raffel2020exploring}. Despite their impressive capabilities, LLMs suffer from hallucination\cite{Radford2019}, often generating confident but incorrect information. This is a critical challenge for knowledge-intensive tasks requiring high accuracy\cite{Radford2019}. Additionally, their static nature being trained on fixed datasets\cite{Radford2019} limits their ability to update knowledge dynamically, making them prone to outdated or inconsistent responses, especially for time-sensitive information.\cite{Mousavi2025}.

Researchers have explored various approaches to address this limitation, including transfer learning\cite{zhuang2020comprehensive}, fine-tuning \cite{Howard2018} which adapt pre-trained models to specific tasks. Despite their strengths, they are computationally expensive and impractical for real-time knowledge updates. A promising alternative is Retrieval-Augmented Generation (RAG) \cite{Chuyuan2025}, which integrates real-time retrieval from external knowledge sources with generative models, enables LLMs to access up-to-date information without retraining,(RAG) faces several limitations. For instance, The performance of RAG heavily depends on the quality of the retriever \cite{karpukhin2020dpr}, as irrelevant or noisy documents can degrade the generator output. Additionally, a critical challenge is determining the optimal number of retrieved passages K to feed into the generator\cite{lewis2020retrieval}. Often resulting in either insufficient or excessive information being retrieved, reduce coherence and increase computational overhead. This trade-off makes K a key factor in balancing performance and efficiency.

@@

\section{Overview of Retrieval-Augmented Generation }\label{sec3}
Retrieval-Augmented Generation (RAG) is a method designed to enhance the performance of large language models (LLMs) by integrating external, reliable knowledge sources into their response generation process\cite{awsRAG}. RAG combines search capabilities with LLM prompting. The process involves feeding both the userss query and the relevant information retrieved by a search algorithm into the LLMs prompt\cite{ilin2023advancedrag}, enabling the model to generate answers based on the provided context.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{rag.JPG}
	\caption{RAG system component\cite{gupta20241}. }\label{rag.PNG}
\end{figure}
As illustrated in figure \ref{rag.PNG}, the RAG system is built around two main components: the retriever and the generator. The retriever's role is to find relevant information from a pre-built data store, while the generator utilizes this retrieved data to produce coherent and contextually appropriate content. The overall RAG process operates as follows :
\subsection{Retrieval in RAG Systems}\label{subsec2}

@@

The query \( q \) represents the user's state at a specific time step, derived from their interaction history. In the MoL (Mixture of Logits) framework, \( q \) is transformed into \( Pq \)
embeddings through a multi-layer perceptron (MLP).
\subsection{Impact of Hyperparameters}
For fair comparison, we maintained consistent architectural choices and training conditions across all experiments, We conducted an extensive hyperparameter analysis comparing both approaches (Hybrid+SAS and MoL+SAS) across different architectural configurations. All experiments were Implemented in TensorFlow and trained on a Google Colab environment with a T4 GPU. We use the Adam optimizer with a learning rate of 0.001. For the hybrid algorithm, we initialize the threshold (Tinit) to 0.3 and adaptively adjust it during training. We discuss detailed hyperparameter settings in table \ref{tab:100k_movies}, table \ref{tab:1m_movies}

\setlength{\tabcolsep}{1pt} % Increase column padding
%
%\begin{table}[ht]
%	\centering
%	\large
%	\renewcommand{\arraystretch}{1.7} % Increase row spacing
%	\resizebox{\textwidth}{!}{  % Scale table to fit width
%		\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
%			\hline
%			\textbf{Model} & \textbf{\shortstack{Max \\ Sequence Length}} & 
%			\textbf{\shortstack{Embedding \\ Dimension}} & 
%			\textbf{\shortstack{Number \\ of Heads}} & 
%			\textbf{\shortstack{Feedforward \\ Dimension}} & 
%			\textbf{\shortstack{Batch \\ Size}} & 
%			\textbf{Epochs} & 
%			\textbf{\shortstack{Val \\ Loss}} & 
%			\textbf{\shortstack{Val \\ Accuracy}} \\
%			\hline
%			Hybrid+SAS & 50  & 128  & 2 & 128  & 128 & 12 & 5.3036 & 0.1584  \\ \hline
%			MoL+SAS    & 50  & 128  & 2 & 128  & 128 & 12 & 5.3152 & 0.1582  \\ \hline
%			Hybrid+SAS & 128 & 256  & 4 & 256  & 128 & 10 & 3.6364 & 0.4301  \\ \hline
%			MoL+SAS    & 128 & 256  & 4 & 256  & 128 & 10 & 3.6374 & 0.4299 \\\hline
%			Hybrid+SAS & 512 & 512  & 4 & 512  & 128 & 10 & 1.2808 & {\bf 0.8120} \\ \hline
%			MoL+SAS    & 512 & 512  & 4 & 512  & 128 & 10 & 1.3050 & 0.8016 \\\hline
%		\end{tabular}
%	}
%	\caption{Results on 100kMovies dataset}
%	\label{tab:100k_movies}
%\end{table}

\begin{table}[ht]
	\centering
	\renewcommand{\arraystretch}{1.5} % Balanced row spacing
	\setlength{\tabcolsep}{0.8em}     % Increased column padding
	\caption{Performance comparison of Hybrid+SAS and MoL+SAS models on the 100kMovies dataset}
	\label{tab:100k_movies}
	\resizebox{\textwidth}{!}{ % Scale to textwidth (remove if too small)
		\begin{tabular}{@{}l*{8}{c}@{}} % Remove vertical lines, add horizontal padding
			\toprule
			\textbf{Model} & 
			\textbf{\makecell{Max \\ Seq. Len.}} & 
			\textbf{\makecell{Embed. \\ Dim.}} & 
			\textbf{Heads} & 
			\textbf{\makecell{FFN \\ Dim.}} & 
			\textbf{\makecell{Batch \\ Size}} & 
			\textbf{Epochs} & 
			\textbf{\makecell{Val. \\ Loss}} & 
			\textbf{\makecell{Val. \\ Acc.}} \\
			\midrule
			Hybrid+SAS & 50  & 128 & 2 & 128 & 128 & 12 & 5.3036 & 0.1584  \\ 
			MoL+SAS    & 50  & 128 & 2 & 128 & 128 & 12 & 5.3152 & 0.1582  \\ 
			\addlinespace[0.2em] % Light spacing between groups
			Hybrid+SAS & 128 & 256 & 4 & 256 & 128 & 10 & 3.6364 & 0.4301  \\ 
			MoL+SAS    & 128 & 256 & 4 & 256 & 128 & 10 & 3.6374 & 0.4299  \\
			\addlinespace[0.2em]
			Hybrid+SAS & 512 & 512 & 4 & 512 & 128 & 10 & \textbf{1.2808} & \textbf{0.8120} \\ 
			MoL+SAS    & 512 & 512 & 4 & 512 & 128 & 10 & 1.3050 & 0.8016  \\
			\bottomrule
		\end{tabular}
	}
\end{table}

@@

%\begin{table}[ht]
%	\centering
%	\large
%	\renewcommand{\arraystretch}{1.7} % Increase row spacing
%	\resizebox{\textwidth}{!}{  % Scale table to fit width
%		\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
%			\hline
%			\textbf{Model} & \textbf{\shortstack{Max \\ Sequence Length}} & 
%			\textbf{\shortstack{Embedding \\ Dimension}} & 
%			\textbf{\shortstack{Number \\ of Heads}} & 
%			\textbf{\shortstack{Feedforward \\ Dimension}} & 
%			\textbf{\shortstack{Batch \\ Size}} & 
%			\textbf{Epochs} & 
%			\textbf{\shortstack{Val \\ Loss}} & 
%			\textbf{\shortstack{Val \\ Accuracy}} \\
%			\hline
%			Hybrid+SAS & 50  & 128  & 2 & 128  & 128 & 10 & 4.5721 & 0.1530  \\ \hline
%			MoL+SAS    & 50  & 128  & 2 & 128  & 128 & 10 & 4.5733 & 0.1526  \\ \hline
%			Hybrid+SAS & 128 & 256  & 4 & 256  & 128 & 10 & 3.4152 & 0.3680  \\ \hline
%			MoL+SAS    & 128 & 256  & 4 & 256  & 128 & 10 & 3.4671 & 0.3627  \\ \hline
%			Hybrid+SAS & 512 & 512  & 4 & 512  & 128 & 10 & 1.0275 & {\bf 0.8350} \\ \hline
%			MoL+SAS    & 512 & 512  & 4 & 512  & 128 & 10 & 1.0350 & 0.8276  \\ \hline
%			
%		\end{tabular}
%	}
%	\caption{Results on 1M Movies dataset}
%	\label{tab:1m_movies}
%\end{table}
\begin{table}[ht]
	\centering
	\renewcommand{\arraystretch}{1.5} % Balanced row spacing
	\setlength{\tabcolsep}{0.8em}    % Improved column padding
	\caption{Performance comparison of Hybrid+SAS and MoL+SAS models on the 1M Movies dataset}
	\label{tab:1m_movies}
	\resizebox{\textwidth}{!}{ % Remove if table appears too compressed
		\begin{tabular}{@{}l*{8}{c}@{}}
			\toprule
			\textbf{Model} & 
			\textbf{\makecell{Max \\ Seq. Len.}} & 
			\textbf{\makecell{Embed. \\ Dim.}} & 
			\textbf{Heads} & 
			\textbf{\makecell{FFN \\ Dim.}} & 
			\textbf{\makecell{Batch \\ Size}} & 
			\textbf{Epochs} & 
			\textbf{\makecell{Val. \\ Loss}} & 
			\textbf{\makecell{Val. \\ Acc.}} \\
			\midrule
			Hybrid+SAS & 50  & 128 & 2 & 128 & 128 & 10 & \textbf{4.5721} & \textbf{0.1530}  \\ 
			MoL+SAS    & 50  & 128 & 2 & 128 & 128 & 10 & 4.5733 & 0.1526  \\ 
			\addlinespace[0.2em]
			Hybrid+SAS & 128 & 256 & 4 & 256 & 128 & 10 & \textbf{3.4152} & \textbf{0.3680}  \\ 
			MoL+SAS    & 128 & 256 & 4 & 256 & 128 & 10 & 3.4671 & 0.3627  \\
			\addlinespace[0.2em]
			Hybrid+SAS & 512 & 512 & 4 & 512 & 128 & 10 & \textbf{1.0275} & \textbf{0.8350} \\ 
			MoL+SAS    & 512 & 512 & 4 & 512 & 128 & 10 & 1.0350 & 0.8276  \\
			\bottomrule
		\end{tabular}
	}
\end{table}

\newpage
Both approaches demonstrate significant performance improvements as model capacity increases, with larger configurations consistently delivering better results. For instance, when increasing the model's capacity such as expanding the (Max Sequence Length: 512, Embedding Dimension: 512, Number of Heads: 4, Feed-Forward Dimension: 512) the hybrid algorithm combined with SASRec (Hybrid+SAS) achieves notable gains, particularly in the most resource-intensive setup. On the ML-100K dataset, Hybrid+SAS reaches a score of\textbf{0.8120} compared to the baseline's \textbf{0.8016}, while on ML-1M, it achieves \textbf{0.8350} versus \textbf{0.8276} The ML-1M dataset generally benefits more from increased capacity, with the performance gap between datasets narrowing as the model scales. While smaller configurations provide a balance of efficiency and performance, the largest configuration, despite its higher computational demands, yields the best results, making it suitable for scenarios where resources are not a constraint. Both approaches show similar benefits from scaling, but Hybrid+SAS maintains a consistent edge in performance.

@@

%
%\section{Research Methods}
%
%\subsection{Part One}
%
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
%malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
%sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
%vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
%lacinia dolor. Integer ultricies commodo sem nec semper.
%
%\subsection{Part Two}
%
%Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
%ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
%ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
%eros. Vivamus non purus placerat, scelerisque diam eu, cursus
%ante. Etiam aliquam tortor auctor efficitur mattis.
%
%\section{Online Resources}
%
%Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
%pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
%enim maximus. Vestibulum gravida massa ut felis suscipit
%congue. Quisque mattis elit a risus ultrices commodo venenatis eget
%dui. Etiam sagittis eleifend elementum.
%
%Nam interdum magna at lectus dignissim, ac dignissim lorem
%rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
%massa et mattis lacinia.
%
%\section{Reproducibility Checklist for JAIR}
%
%Select the answers that apply to your research -- one per item. 
%
%\subsection*{All articles:}
%
%%\hh{revised for stylistic consistency:}
%\begin{enumerate}
